---
title: "Coursera Statistical Inference Project - Part 1"
author: "Father Abraham"
date: "13/09/2014"
output: pdf_document
---

**Part 1**

The first task of the project is to "[i]llustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s".

```{r}
##set seed for predictable random numbers (!)
set.seed(23)

##satisfy project requirements
sims=1000
lambda=0.2
obs=40

##load and format data
dat1000 <- as.data.frame(replicate(sims, mean(rexp(obs, lambda))))
names(dat1000) <- c("Means")
```

With the data in place we will now plot the distribution as a histogram using the following code, and we add a line to show where the distribution is centered - very close to 5.:

```{r echo=FALSE}
library(ggplot2)
library(ggthemes)
g <- ggplot(dat1000, aes(x=Means)) + geom_histogram(aes(y=..density..), binwidth=.25,  colour="black", fill="white")
g <- g + ggtitle("Average of Distributions")
g <- g + xlab("Mean") + ylab("Density")
g <- g + theme_economist()
g <- g + geom_vline(aes(xintercept=mean(dat$Means)), color="blue", linetype="dashed", size=1)
g
```

Let's have a look at some other properties of the original distribution.  

The variance:

```{r}
var(dat1000$Means)
```

The theoretical variance is:

???

The standard deviation:

```{r}
sqrt(var(dat1000$Means))
```


Next, we plot the normal distribution on top, we know the mean is 5 from 1/lambda, so we will use that (i.e. we won't *fully* standardize, by centering the mean on 0), and for a standard normal distribution the standard deviation is 1 (Source: http://en.wikipedia.org/wiki/Normal_distribution) 

```{r}
g <- g + stat_function(fun = dnorm, colour = "red", arg = list(mean=5, sd =1))
g
```


Clearly, the distribution is very close to a normal distribution, and we have demonstrated that the Central Limit Theorem is not refuted here.

I don't yet know how to create the 95% Confidence Level....

